---
title: "02-exercises"
author: "Christopher Brown"
date: "April 13, 2016"
output: html_document
---

This exercise uses the **Fuel Economy** data set from the **AppliedPredicitiveModeling** package.

Note: The following will set-up your environment for this exercise. If you get an error stating that the packages have not been found, you need to install those packages.


```{r,echo=FALSE}

packs <-  c('AppliedPredictiveModeling', 'ggplot2', 'magrittr', 'dplyr')

for( nm in packs ) { 
  # message(nm)
  if( ! nm  %in% installed.packages()[,1]  ) install.packages(nm)
  library(nm, character.only = TRUE)
}

# Load data set into environment
data(FuelEconomy)
... = NULL  # Needed for aesthetics 

```



## Exercise 1

Hint: See `?cars2010`


- After the **Fuel Economy** data is loaded, combine three data sets into one data set. (Note: The name `dat` is very often used in these situations, `data` is a reserved R word.)  

```{r}
dat <- bind_rows( cars2010, cars2011, cars2012)    # Define 
```


- What is a good "naive guess" of FE? Show your work

```{r}
naive_guess <- mean(dat$FE)
```


- plot FE (Fuel Econonomy) vs. EngDisp. Plot the naive guess.

```{r}
qplot( y=dat$FE, x=dat$EngDispl ) + geom_hline(yintercept=naive_guess)
```

- Sample 10 observations from `dat`
- Plot this data. Add a line for the naive_guess. 


```{r}

set.seed(314) 

# Sample 
samp <- dat %>% dplyr::sample_n(10) 
samp_naive_guess <- mean(samp$FE)

# Plots 
qplot( y=samp$FE, x=samp$EngDispl ) + geom_hline(yintercept=samp_naive_guess)


```


## Exercise 2:


Write a loss functions for calculating:
- Root Mean Square Error
- Mean Absolute Error
- Median Absolute Error

All functions should accept two arguments: 


```{r}

rmse <- function(y,yhat) {
  ( y - yhat )^2  %>% mean %>% sqrt 
}

mae <- function(y, yhat) {
  ( y - yhat )  %>% abs %>% mean
}

medae <- function(y, yhat) { 
  ( y - yhat )  %>% abs %>% median
}

```

Use these functions to evaluate the loss/performance of:
- the naive guess


```{r}

rmse(dat$FE, naive_guess)
mae(dat$FE, naive_guess)
medae(dat$FE, naive_guess)

```


## Exercise 3: Linear Model and Model Performance 

- Use `lm` to create a linear model fitting the relationship between FE and EngDispl for the cars2010 data set


```{r}
fit_2010 <- lm(FE ~ EngDispl, data = cars2010) 
```

- Use your functions to evaluate the training error

```{r}
y_2010 <- predict( fit_2010, data=cars2010 ) 
rmse_2010 <- rmse(cars2010$FE, y_2010)

mae(cars2010$FE, y_2010)
medae(cars2010$FE, y_2010)
```


- Use your model to:
-- predict the FE for 2011. What is the RMSE errors associated with the predictions.
-- predict the FE for 2012. What is the RMSE errors associated with the predictions.

```{r}
y_2011 <- predict( fit_2010, data=cars2011 ) 
y_2012 <- predict( fit_2010, data=cars2012 )

rmse_2011 <- rmse(cars2011$FE, y_2011)
rmse_2012 <- rmse(cars2012$FE, y_2012)


```


```{r}
# DO NOT EDIT 
rmse_2010 
rmse_2011
rmse_2012
```


## Exercise 4: 

- Model the fuel economy (`FE`) as a function of `EngDispl`, `NumCyl` and `VarValve` using the `cars2011` data set.  
- Provide betas


```{r}
fit_more_2011 <- lm(FE ~ EngDispl + NumCyl + VarValveTiming + VarValveLift, data = cars2011) 
betas <- summary(fit_more_2011)$coefficient[,1]
betas
```
