---
title: "03-exercises"
author: "Phil Li"
date: "April 20, 2016"
output: html_document
---

## Readings

***APM***

- Chapter 4 "Over Fitting and Model Tuning"
- Chapter 12.2 "Logistic Regression""


## Miscellaneous

I am still struggling with names ...

- Please send me your picture
- [my photo here!](https://media.licdn.com/mpr/mpr/shrinknp_400_400/p/5/000/229/29f/3b94812.jpg)


## Assignment 

Note: The following will set-up your environment for this exercise. If you get an error stating that the packages have not been found, you need to install those packages.


```{r,echo=FALSE, warning=FALSE, message=FALSE}

packs <-  c('AppliedPredictiveModeling', 'ggplot2', 'magrittr', 'dplyr', 'caret')

for( nm in packs ) { 
  # message(nm)
  if( ! nm  %in% installed.packages()[,1]  ) install.packages(nm)
  library(nm, character.only = TRUE)
}

# Load data set into environment
data(FuelEconomy)     # See ?cars2010
fe <- dplyr::bind_rows(cars2010, cars2011, cars2012)    # Define Da


data("GermanCredit")  # see GermanCredit

... = NULL  # Needed for aesthetics 

```


## StepAIC


Using Fuel Economy data set from the **AppliedPredictiveModeling** Package.
- fit the simplest possible model using lm

```{r}

 # this is essentially the mean
library(MASS)
fit_simplest <- lm(FE ~ 1,fe)
```


- Use MASS::StepAIC to improve the model using forward stepwise regression

```{r}

# take the entire list of variables and remove the output variable to get the input variables... 
fe %>% names %>% setdiff("FE") %>% paste(collapse=" + ") -> dependent_variables
stepAIC_formula <- paste("FE ~ ", dependent_variables)

# stick the input variables into forward AIC
# AIC fwd works by adding the input variable that reduces AIC most, then repeating that until no further AIC reduction is possible
# (AIC is a measure of the tradeoff between model perfomance and complexity.)
step_forward <- stepAIC ( fit_simplest, scope = stepAIC_formula, direction = "forward")

```


- Fit the "full" model using lm
- USe MASS::StepAIC to improve the model using backward stepwise regression 

```{r}

# AIC backwards is the opposite... put in all variables first... take one out at a time 
fit_full <- lm(FE ~ ., fe)
#stepAIC(fit_full, score = ~1, direction = "backward")
step_backward <- stepAIC(fit_full, scope = ~1, direction = "backward")


```

- Are they the same model? If not why?  Which is better?
```{r}
# The two models are quite similar. Forward has 9 input variables.  Backward has 12.  8 of those are in common.
step_forward$call
step_backward$call

# The two models have almost identical parsimony measures, as revealed in the output earlier....
# forward AIC=3662.66
# backward AIC=3663.23

# The R-squares are nearly identical also.

#forward StepAIC result 
summary(lm(formula = FE ~ EngDispl + CarlineClassDesc + DriveDesc + Transmission + 
    NumCyl + IntakeValvePerCyl + VarValveLift + TransLockup + 
    NumGears, data = fe))$r.squared

#backward StepAIC result
summary(lm(formula = FE ~ EngDispl + NumCyl + Transmission + AirAspirationMethod + 
    NumGears + TransLockup + TransCreeperGear + DriveDesc + ExhaustValvesPerCyl + 
    CarlineClassDesc + VarValveLift, data = fe))$r.squared

# If simplicity is desired, forward seems preferred as it has fewer input variables. 

```




## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`


```{r}

logistic <- function(x) { 
  1 / (1 + exp(1)^(-x) ) 
}

curve(logistic, from=-10, to=10)

logistic_inv <- function(y) { 
  - (log((1-y)/y))
}

curve(logistic_inv, from=0, to=1)

```

**NOTE"** These functions are quite handy, in evaluating logistic regression results. You may want to save these functions in your own package.  

```{r}
# DO NOT EDIT
c(-Inf,0,Inf) %>% logistic

c(0,0.5,1) %>% logistic_inv

```


## German Credit Model

Using the GermanCredit data from the **Caret** package/ UCI Machine Learning Library, create a model for `Class` ("Good" vs. "Bad" ). Show your model performance.  

```{r}
#data(GermandCredit) #didn't we already call this?

# take the entire list of variables and remove the output variable to get the input variables... 
GermanCredit %>% names %>% setdiff("Class") %>% paste(collapse=" + ") -> GermanCredit_dependent_variables
GermanCredit_stepAIC_formula <- paste("Class ~ ", GermanCredit_dependent_variables)

GermanCredit_fit_simplest <- glm(Class ~ 1, data=GermanCredit, family="binomial")
german_credit_step_forward <- stepAIC ( GermanCredit_fit_simplest, scope = GermanCredit_stepAIC_formula, direction = "forward")

german_credit_fit_full <- glm(Class ~ ., data=GermanCredit, family="binomial")
german_credit_step_backward <- stepAIC(german_credit_fit_full, scope = ~1, direction = "backward")

# best model gets a prize!
# AIC can also use stepAIC
```



## Iterative Correlated Feature Removal 

- Implement Kuhn's iterative feature removal function described in **APM** Section 3.5, page 47


```{r}
# 1. Calculate the correlation matrix of the predictors.
# 2. Determine the two predictors associated with the largest absolute pairwise correlation (call them predictors A and B).
# 3. Determine the average correlation between A and the other variables. Do the same for predictor B.
# 4. If A has a larger average correlation, remove it; otherwise, remove predictor B.
# 5. Repeat Steps 2–4 until no absolute correlations are above the threshold.

Kuhns_removal <- function(myData) {
  cor_matrix <- cor(myData[sapply(myData, is.numeric)])  # step 1
  
}

Kuhns_removal(fe)

#cor(x=fe, y=fe, use="pairwise.complete.obs", na.rm = TRUE, method="kendall") 
```




## Synthetic Data (Optional)

Sometimes it is useful to "synthesize" feature data for to understand how a certain model behaves. 
Sythesize the following features 1000-element vectors: 

- x1: a normally distributed variable with `mean = 20` and standard deviation = 20 (`sd=8`).
- x2: a log-normally distributed feature with `meanlog = 1`, `sdlog=1.2`
- x3: a uniformly distributed feature with `min=0` and `max=50`. 

```{r}
nsamples = 20

x1 <- rnorm(nsamples,20,20)  
x2 <- rlnorm(nsamples, meanlog=1, sdlog = 1.2)
x3 <- runif(nsamples,0,50)

```

Next synthesis a response, `y` using the betas provided and an intercept that is normally distributed at 20 with standard deviation of 2. (**Hint:**  The betas thought of can be a vector or matrix)



```{r}

beta0 <- rnorm(nsamples,0,15)  # intercept!
beta1 <- 2.3
beta2 <- 4
beta3 <- 7

betas <- matrix( c(2.5, 4, 7), nrow=1  )  # 1x4 matrix

# x0 <- rep(1,nsamples) 

X  <- cbind(x1,x2,x3)  # 1000x4

y <- betas %*% t(X) %>% t
y <- y + beta0

qplot(y)
dat <- data.frame(y,X)

fit <- lm( y ~ . , dat )

coef(fit)

fit
```

- Did you recover the betas? 
- Is the model good?
- What happens if increase the value of `nsamples`? Decrease it?
- What transformations would you apply to x1? x2? x3? 

